---
title: "Reproduciable Research Assignement 1"
output: html_document
---

Analze personal movement activity generated by device like Fitbit, Nike, etc.  Here are the analysis that will be perform in this research.
1. What is the mean total number of steps taken per day?
2. What is teh daily acitivity pattern?
3. Imputing missing values.
4. Are there diferences in the activity patterns between weekdays and weekends?

Let's Begin
Make sure you set the working directory to your data source file location. 

Loading source file and load library(dplyr) package
```{r, echo=TRUE}

activity <- read.csv("activity.csv")
library(dplyr)

```

1. Calcualte mean total number of steps taken per day:

```{r, echo=TRUE}

GroupByDay <- group_by(activity, date)
TotalStepsPerDay<- summarize(GroupByDay, TotalSteps = sum(steps, na.rm = TRUE))
hist(TotalStepsPerDay$TotalSteps)
mean(TotalStepsPerDay$TotalSteps)
median(TotalStepsPerDay$TotalSteps)
```

2. What is the daily activity pattern
```{r, echo=TRUE}

GroupByInterval <- group_by(activity, interval)

## Plot average steps by time interval
AvgStepsByInterval<- summarize(GroupByInterval, AvgSteps = mean(steps, na.rm = TRUE))
plot(AvgStepsByInterval$interval, AvgStepsByInterval$AvgSteps, type="l", xlab = "5 minute daily interval", ylab = "Average Steps")

## Display interval with the average maximum steps
filter(AvgStepsByInterval, AvgSteps == max(AvgStepsByInterval$AvgSteps))

```

3. Imputing missing values
```{r, echo=TRUE}

## Count of number of NAs in teh dataset
sum(is.na(activity$steps))

## Crate a dataset and fill teh missing values with average for the interval
## Use the average step calculate in question 2 
actvityWithMean <- inner_join(activity, AvgStepsByInterval, by = NULL)
actvityWithMean$steps[is.na(actvityWithMean$steps)] <- actvityWithMean$AvgSteps

GroupByDayNoNA <- group_by(actvityWithMean, date)
TotalStepsPerDayNoNA<- summarize(GroupByDayNoNA, TotalSteps = sum(steps, na.rm = TRUE))
hist(TotalStepsPerDayNoNA$TotalSteps)
mean(TotalStepsPerDayNoNA$TotalSteps)
median(TotalStepsPerDayNoNA$TotalSteps)

```
