---
title: "Reproduciable Research Assignement 1"
output: html_document
---

Analze personal movement activity generated by device like Fitbit, Nike, etc.  Here are the analysis that will be perform in this research.
1. What is the mean total number of steps taken per day?
2. What is the daily acitivity pattern?
3. Imputing missing values.
4. Are there diferences in the activity patterns between weekdays and weekends?

Let's Begin
Make sure you set the working directory to your data source file location. 

Loading source file and load library(dplyr) package
```{r, echo=TRUE}

activity <- read.csv("activity.csv")
library(dplyr)

```

1. Calcualte mean total number of steps taken per day:

```{r, echo=TRUE}

GroupByDay <- group_by(activity, date)
TotalStepsPerDay<- summarize(GroupByDay, TotalSteps = sum(steps, na.rm = TRUE))
hist(TotalStepsPerDay$TotalSteps)
mean(TotalStepsPerDay$TotalSteps)
median(TotalStepsPerDay$TotalSteps)

```

2. What is the daily activity pattern
```{r, echo=TRUE}

GroupByInterval <- group_by(activity, interval)

## Plot average steps by time interval
AvgStepsByInterval<- summarize(GroupByInterval, AvgSteps = mean(steps, na.rm = TRUE))
plot(AvgStepsByInterval$interval, AvgStepsByInterval$AvgSteps, type="l", xlab = "5 minute daily interval", ylab = "Average Steps")

## Display interval with the average maximum steps
filter(AvgStepsByInterval, AvgSteps == max(AvgStepsByInterval$AvgSteps))

```

3. Imputing missing values
```{r, echo=TRUE}

## Count of number of NAs in the dataset
sum(is.na(activity$steps))

## Crate a dataset and fill teh missing values with average for the interval
## Use the average step calculate in question 2 
actvityWithMean <- inner_join(activity, AvgStepsByInterval, by = NULL)
actvityWithMean$steps[is.na(actvityWithMean$steps)] <- actvityWithMean$AvgSteps

##Display histogram with total steps without NA
GroupByDayNoNA <- group_by(actvityWithMean, date)
TotalStepsPerDayNoNA<- summarize(GroupByDayNoNA, TotalSteps = sum(steps, na.rm = TRUE))
hist(TotalStepsPerDayNoNA$TotalSteps)
## Mean value
mean(TotalStepsPerDayNoNA$TotalSteps)
## Median value
median(TotalStepsPerDayNoNA$TotalSteps)

```

4. Are there diference between weekdays and weekends?

```{r, echo=TRUE}
## Create a new column with factor fo Weekday and Weekend
actvityWithMean <- mutate(actvityWithMean, DateCategory = weekdays(as.Date(actvityWithMean$date)))
actvityWithMean$DateCategory <- factor(actvityWithMean$DateCategory)
levels(actvityWithMean$DateCategory) <- list(Weekday = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), Weekend = c("Saturday", "Sunday"))

## group average step by interval and datecategory(Weekday and weekend)
GroupByIntervalwithFactor <- group_by(actvityWithMean, interval, DateCategory)
AvgStepsByIntervalwithFactor <- summarize(GroupByIntervalwithFactor, AvgSteps = mean(steps, na.rm = TRUE))

## Filter out dataset by weekday or weekend
JustWeekDay <- filter(AvgStepsByIntervalwithFactor, DateCategory == "Weekday")
JustWeekEnd <-filter(AvgStepsByIntervalwithFactor, DateCategory == "Weekend")

## Make the plot
par(mfrow = c(2, 1))
plot(JustWeekDay$interval, JustWeekDay$AvgSteps, type="l", xlab = "5 minute daily interval", ylab = "Average Steps", main = "Weekday")
plot(JustWeekEnd$interval, JustWeekEnd$AvgSteps, type="l", xlab = "5 minute daily interval", ylab = "Average Steps", main = "Weekend")

```

